---
title: "Moody Analysis"
Name: Revanth Korrapolu (rrk69)
---


```{r}
Moody <- read.csv("MOODY-2019.csv")
```


Format data so that studentid is the index
```{r}
Moody2 <- Moody[-1]
row.names(Moody2) <- Moody$STUDENTID
```



Create Dummy variable for On_Smartphone, Asks_Questions, Leaves_Early, Late_In_Class,
(convert categorical data into continuous)
```{r}
Moody2$dummy_on_smartphone[Moody$ON_SMARTPHONE == "always"] <- "4"
Moody2$dummy_on_smartphone[Moody$ON_SMARTPHONE == "frequently"] <- "3"
Moody2$dummy_on_smartphone[Moody$ON_SMARTPHONE == "rarely"] <- "2"
Moody2$dummy_on_smartphone[Moody$ON_SMARTPHONE == "never"] <- "1"
Moody2$dummy_on_smartphone <- factor(Moody2$dummy_on_smartphone)

Moody2$dummy_grade[Moody$GRADE == "A"] <- "5"
Moody2$dummy_grade[Moody$GRADE == "B"] <- "4"
Moody2$dummy_grade[Moody$GRADE == "C"] <- "3"
Moody2$dummy_grade[Moody$GRADE == "D"] <- "2"
Moody2$dummy_grade[Moody$GRADE == "F"] <- "1"
Moody2$dummy_grade <- factor(Moody2$dummy_grade)

Moody2$dummy_asks_questions[Moody$ASKS_QUESTIONS == "always"] <- "3"
Moody2$dummy_asks_questions[Moody$ASKS_QUESTIONS == "sometimes"] <- "2"
Moody2$dummy_asks_questions[Moody$ASKS_QUESTIONS == "never"] <- "1"
Moody2$dummy_asks_questions <- factor(Moody2$dummy_asks_questions)

Moody2$dummy_leaves_early[Moody$LEAVES_EARLY == "always"] <- "3"
Moody2$dummy_leaves_early[Moody$LEAVES_EARLY == "rarely"] <- "2"
Moody2$dummy_leaves_early[Moody$LEAVES_EARLY == "never"] <- "1"
Moody2$dummy_leaves_early <- factor(Moody2$dummy_leaves_early)

Moody2$dummy_late_to_class[Moody$LATE_IN_CLASS == "Yes"] <- "3"
Moody2$dummy_late_to_class[Moody$LATE_IN_CLASS == "Sometimes"] <- "2"
Moody2$dummy_late_to_class[Moody$LATE_IN_CLASS == "Never"] <- "1"
Moody2$dummy_late_to_class <- factor(Moody2$dummy_late_to_class)
```


Sort Moody data by dummy variables

```{r}
Moody2 <- Moody2[order(-Moody2[,1], Moody2[,7], Moody2[,8], Moody2[,9], Moody2[,10], Moody2[,11]),]
```


Convert dummy variables into regression variables

```{r}
for(i in 1:nrow(Moody2)) {
    n <- 0    
    if (Moody2[i,7] == 1){
       n <- 1
    }
    Moody2$smartphone1[i] <- n
    
    n <- 0    
    if (Moody2[i,7] == 2){
       n <- 1
    }
    Moody2$smartphone2[i] <- n
    
    n <- 0    
    if (Moody2[i,7] == 3){
       n <- 1
    }
    Moody2$smartphone3[i] <- n
    
    n <- 0    
    if (Moody2[i,9] == 1){
       n <- 1
    }
    Moody2$question1[i] <- n
    
    n <- 0    
    if (Moody2[i,9] == 2){
       n <- 1
    }
    Moody2$question2[i] <- n
    
    n <- 0    
    if (Moody2[i,9] == 3){
       n <- 1
    }
    Moody2$question3[i] <- n
    
    n <- 0    
    if (Moody2[i,10] == 1){
       n <- 1
    }
    Moody2$leaves1[i] <- n
    
    n <- 0    
    if (Moody2[i,10] == 2){
       n <- 1
    }
    Moody2$leaves2[i] <- n
    
    n <- 0    
    if (Moody2[i,10] == 3){
       n <- 1
    }
    Moody2$leaves3[i] <- n
    
    n <- 0    
    if (Moody2[i,11] == 1){
       n <- 1
    }
    Moody2$late1[i] <- n
    
        n <- 0    
    if (Moody2[i,11] == 2){
       n <- 1
    }
    Moody2$late2[i] <- n
    
        n <- 0    
    if (Moody2[i,11] == 3){
       n <- 1
    }
    Moody2$late3[i] <- n
    
}
```

For calculating linear regression

```{r}
fit <- lm(as.numeric(Moody2$dummy_grade) ~ Moody2$SCORE + Moody2$smartphone1 + Moody2$smartphone2 + Moody2$smartphone3 + Moody2$question1 + Moody2$question2 + Moody2$question3 + Moody2$leaves1 + Moody2$leaves2 + Moody2$leaves3 + Moody2$late1 + Moody2$late2 + Moody2$late3)

summary(fit)
```

There were many colinear variables, however, we were still able to identify the most signficant variables as well as the coefficients.  


These scatter plots show Grade vs Score based on the various categorical variables.

```{r}
library(ggplot2)

ggplot(Moody2, aes(GRADE, SCORE, colour = ON_SMARTPHONE)) + geom_point(position = position_dodge(width = 0.4)) + ggtitle("On Smartphone")

ggplot(Moody2, aes(GRADE, SCORE, colour = LATE_IN_CLASS)) + geom_point(position = position_dodge(width = 0.4)) + ggtitle("Late In Class")

ggplot(Moody2, aes(GRADE, SCORE, colour = ASKS_QUESTIONS)) + geom_point(position = position_dodge(width = 0.4)) + ggtitle("Asks Questions")

ggplot(Moody2, aes(GRADE, SCORE, colour = LEAVES_EARLY)) + geom_point(position = position_dodge(width = 0.4)) + ggtitle("Leaves Early")
```



 

